Query Optimization for Dynamic Imputation

### 2017-05-15 paper/experiments.tex
paper: minor changes

We evaluate \ProjectName{} optimizing for quality ($\alpha=0$), optimizing for performance
($\alpha=1$), and targeting a balance of each. As a baseline, we fully impute the tables
used by each query and then run the query on the imputed tables, simulating the process that
an analyst would follow to apply existing imputation techniques.

\Cref{fig:runtimes} shows a summary of the performance results. The quality-optimized
queries are an order-of-magnitude faster than the baseline. We get another order-of-magnitude speedup when optimizing for performance. This performance
differential means it is feasible
to explore multiple imputation operations (including more expensive operators) when using
\ProjectName{}, in contrast to the traditional approach of fully imputing the dataset.

@@

\review{
We also experiment with mean-value imputation and hot
deck imputation, as outlined in \Cref{sec:imputation}. One approach to
mean-value imputation, to facilitate non-blocking operation within the iterator
model, is to estimate the column mean from base table histograms. In this case,
imputations are identical no matter where the operator is placed in the query
plan. Moreover, runtimes are essentially identical regardless of the value of
$\alpha$ because the time cost of the imputation is relatively low no matter
how many tuples need to be imputed. In hot deck imputation, runtimes are also
similar across $\alpha$, because the cost of buffering
tuples and sampling at random is negligible compared to the rest of query
execution. \todo{be quantitative: In terms of accuracy, relative errors using hot deck are close to
zero; since the baseline, in this case, is hot deck imputation on the base
table, only skew in the distributions arising from filters and joins would lead
to larger errors. In some cases, skew does arise, and errors are on the same
magnitude as in the decision trees case. Indeed, \ProjectName{} yields larger
benefit when using relatively sophisticated, higher quality imputation
algorithms as opposed to simpler strategies that would even be quick to
implement on the base table.
}
}

@@

Planning times for our benchmark queries are reasonable, with mean planning times across the
benchmark queries ranging from \planningtimelow{} to \planningtimehigh{}, and ranging from
\planningtimelowpct{} to \planningtimehighpct{}\% of total runtime. In 99\% of individual
cases, the optimizer returned a query plan with \planningtimepninetynine{}. The choice of
$\alpha$ has no impact on planning times, because query selection is done after collecting
the Pareto frontier.

>>>>

We evaluate \ProjectName{} optimizing for quality ($\alpha=0$), optimizing for performance
($\alpha=1$), and targeting a balance of each ($\alpha = 0.5$). As a baseline, we fully impute the tables
used by each query and then run the query on the imputed tables, simulating the process that
an analyst would follow to apply existing imputation techniques.

\Cref{fig:runtimes} summarizes the performance results. The quality-optimized
queries are an order-of-magnitude faster than the baseline. We get another order-of-magnitude speedup when optimizing for performance. This performance
differential means it is feasible
to explore multiple imputation operations (including more expensive operators) when using
\ProjectName{}, in contrast to the traditional approach of fully imputing the dataset.

@@

\review{
We also experiment with mean-value imputation and hot
deck imputation (\Cref{sec:imputation}). One approach to
mean-value imputation, to facilitate non-blocking operation within the iterator
model, is to estimate the column mean from base table histograms. In this case,
imputations are identical no matter where the operator is placed in the query
plan. Moreover, runtimes are essentially identical regardless of the value of
$\alpha$ because the time cost of the imputation is relatively low no matter
how many tuples need to be imputed. In hot deck imputation, runtimes are also
similar across $\alpha$, because the cost of buffering
tuples and sampling at random is negligible compared to the rest of query
execution. \todo{be quantitative: In terms of accuracy, relative errors using hot deck are close to
zero; since the baseline, in this case, is hot deck imputation on the base
table, only skew in the distributions arising from filters and joins would lead
to larger errors. In some cases, skew does arise, and errors are on the same
magnitude as in the decision trees case. Indeed, \ProjectName{} yields larger
benefit when using relatively sophisticated, higher quality imputation
algorithms as opposed to simpler strategies that would even be quick to
implement on the base table.
}
}

@@

Planning times for our benchmark queries are reasonable, with mean planning times across the
benchmark queries ranging from \planningtimelow{} to \planningtimehigh{}, and ranging from
\planningtimelowpct{} to \planningtimehighpct{}\% of total runtime. In 99\% of individual
cases, the optimizer returned a query plan within \planningtimepninetynine{}. The choice of
$\alpha$ has no impact on planning times, because query selection is done after collecting
the Pareto frontier.
